Scaffolder - Software for \change{Reproducible}{Manual} Genome Scaffolding.

### 2011-08-15 src/article.tex
Revise results section

The Scaffolder software facilitates reproducibly joining nucleotide sequences
together into a single contiguous scaffolded super-sequence. Plain-text
scaffold files written in the YAML syntax specify how these sequences should
be joined and the scaffolder software is used to generate the scaffold
sequence from these instructions. In addition to specifying which contigs are
required, the scaffold file allows the contigs to be edited into smaller
sub-sequences or reverse complemented if necessary. The process of genome
finishing may involve producing additional oligonucleotide sequences to fill
unknown regions in the scaffold and the Scaffolder allows these additional
insert sequences to be used to fill these gaps. These inserts can also be
treated in the same manner as larger contig sequences: trimmed and/or reverse
complemented to match the corresponding gap region size and orientation. \pb

The distances between contigs may be estimated from paired-read data or from
mapping the contigs to a reference genome. These inter-contig gap regions are
useful to join separate sequences together by an estimated size and the
scaffold file allows for the specification of such unresolved regions by
inserting regions of `N' nucleotides into the scaffold. The use of these
regions in the scaffold indicates the unresolved regions in the build and their
approximate size. \pb

The scaffold file specifies how sequence regions are joined together; however,
the sequences themselves are not stored in the scaffold file and are instead
maintained as a separate FASTA file. Nucleotide sequences in a FASTA-format
file are the standard output for a genome assembler, and these may be contigs,
sets of contigs scaffolded into larger sequences, or both. These nucleotide
sequences stored in the FASTA file are referenced in the scaffold using the
first word from in the FASTA header of the corresponding sequence. Maintaining
the nucleotide sequences in a separate file preserves the unedited sequence
and decouples the data from the specification of how it should be used to
produce the genome sequence. \pb

@@

The scaffold file is written using the YAML syntax and an example is shown in
Figure 1. This file illustrates the text attributes used to describe
a scaffold and how the sequences are correspondingly joined together in the
genome build. The basic layout of the scaffold file is a list of entries,
where each entry corresponds to a region of sequence in the generated scaffold
super-sequence. \add{Each scaffold file represents one scaffolded nucleotide
sequence and separate scaffolds are defined in separate files.} \pb

@@

The last entry in the scaffold, highlighted in blue, adds the sequence named
`sequence2' to the scaffold. This entry demonstrates how this sequence may be
manipulated prior to addition to the scaffold. The `start' and `stop' tags trim
the sequence to these coordinates inclusively, while the ``reverse'' tag
instructs scaffolder to also reverse complement the sequence. In the putative
scaffold shown in Figure 1 this completes the scaffold sequence. \pb

The final entry in the scaffold also includes the ``inserts'' tag to add
additional regions of sequence. These inserts are also added as a YAML list,
with each insert beginning with a dash. The first insert, shown in purple,
uses similar attributes to that of a sequence entry; the reverse, start and
stop tags are used to trim and reverse complement the insert. Similarly the
`source' tag identifies the corresponding FASTA sequence as `insert1'. The
``open'' and ``close'' tags are specific to inserts and determine where the
insert is added in the enclosing sequence. The region of the sequence inside
these coordinates is inclusively replaced by the specified insert sequence.
This is visualised in the putative scaffold in Figure 1 by the black lines
bisecting the blue sequence. \pb

@@

Scaffolder provides a standardised set of Ruby classes and methods (API) for
interacting with the scaffold. This allows Scaffolder to be integrated into
existing genomics workflows or used with Ruby build tools such as Rake. In
addition Scaffolder provides a command line interface (CLI) to validate the
scaffold file and build the draft super sequence. The Scaffolder CLI behaves
as a standard Unix tool and returns appropriate exit codes and manual pages.
The use of both these Scaffolder interfaces is outlined in detail on the
scaffolder website (\scaffolder). \add{This website also contains a ``getting
started'' guide which introduces using Scaffolder.} \pb

>>>>

The Scaffolder software facilitates reproducibly joining nucleotide sequences
together into a single contiguous scaffolded super-sequence. Plain-text
scaffold files written in \remove{the} YAML \remove{syntax} specify how these
sequences should be joined\change{. and the}{The} scaffolder software is used
to generate the scaffold sequence from these instructions. In addition to
specifying which contigs are required, the scaffold file allows the contigs to
be edited into smaller sub-sequences or reverse complemented if necessary.\pb

The process of genome finishing may involve producing additional
oligonucleotide sequences to fill unknown regions in \change{the}{a} scaffold.
\remove{and} The Scaffolder \change{allows}{format provides for using} these
additional insert sequences to \remove{be used} to fill \remove{these} gaps.
These inserts can also be treated in the same manner as larger contig
sequences: trimmed and/or reverse complemented to match the corresponding gap
region size and orientation. \pb

The distances between contigs may be estimated from paired-read data or from
mapping the contigs to a reference genome. These inter-contig gap regions are
useful to join separate sequences together by \change{an}{the} estimated
\change{size and }{distance.} The scaffold file allows for the specification
of such unresolved regions by inserting regions of `N' nucleotides into the
scaffold. The use of these regions in the scaffold indicates the unresolved
regions in the build and their approximate size. \pb

\change{The scaffold file specifies how sequence regions are joined together;
however, the sequences themselves are no  stored in the scaffold file and are
instead}{The nucleotide sequences used in the scaffold are} maintained as
a separate FASTA file. \remove{Nucleotide sequences in a FASTA-format file are
the standard output for a genome assembler, and these may be contigs, sets of
contigs scaffolded into larger sequences, or both. }The nucleotide sequences
\remove{stored in the FASTA file} are referenced in the scaffold using the
first word from in the FASTA header of the corresponding sequence. Maintaining
the nucleotide sequences in a separate file preserves the unedited sequence
and decouples the data from the specification of how it should be used to
produce the genome sequence. \pb

@@

The scaffold file is written using the YAML syntax and an example is shown in
Figure 1. This file illustrates the text attributes used to describe
a scaffold and how the sequences are correspondingly joined together in the
genome build. The basic layout of the scaffold file is a list of entries,
where each entry corresponds to a region of sequence in the generated scaffold
super-sequence. \add{Each scaffold file represents one scaffolded nucleotide
sequence and separate scaffolds should be defined in separate files.} \pb

@@

The last entry in the scaffold, highlighted in blue, adds the sequence named
`sequence2' to the scaffold. This entry demonstrates how this sequence may be
manipulated prior to addition to the scaffold. The `start' and `stop' tags
trim the sequence to these coordinates inclusively.\remove{while} The
``reverse'' tag instructs scaffolder to also reverse complement the sequence.
In the putative scaffold shown in Figure 1 this completes the \remove{scaffold
} sequence. \pb

The final entry in the scaffold \remove{also} includes the ``inserts'' tag to
add additional regions of sequence. These inserts are also added as a YAML
list, with each insert beginning with a dash. The first insert, shown in
purple, uses similar attributes to that of a sequence entry; the reverse,
start and stop tags are used to trim and reverse complement the insert.
Similarly the `source' tag identifies the corresponding FASTA sequence as
`insert1'. The ``open'' and ``close'' tags are specific to inserts and
determine where the insert is added in the enclosing sequence. The region of
the sequence inside these coordinates is inclusively replaced by the specified
insert sequence. This is visualised in the putative scaffold in Figure 1 by
the black lines bisecting the blue sequence. \pb

@@

Scaffolder provides a standardised set of Ruby classes and methods (API) for
interacting with the scaffold. This allows Scaffolder to be integrated into
existing genomics workflows or used with Ruby build tools such as Rake. In
addition Scaffolder provides a command line interface (CLI) to validate the
scaffold file and build the draft super sequence. The Scaffolder CLI behaves
as a standard Unix tool and returns appropriate exit codes and manual pages.
The use of both these Scaffolder interfaces is outlined in detail on the
scaffolder website (\scaffolder). \add{This website provides a ``getting
started'' guide as an introduction to using Scaffolder to build a genome
scaffold.} \pb
